---
title: "viz_i"
output: github_document
date: "2022-09-28"
---

```{r setup}
library(tidyverse)
library(ggridges)
library(patchwork)
```
## load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(  ##access to the public online data rather than read a file
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(  ##reorganize the dataset
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Scatterplots
Create my first scatterplot ever

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()

```

new approach, same plot:

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```
save and edit a plot object

```{r}
weather_plot =
  weather_df %>%
  ggplot(aes(x = tmin,y = tmax))

weather_plot + geom_point()
  
```

## Advancedd scatterplot

start with the same one and make it fancy!

```{r}
weather_df %>%
  ggplot(aes(x = tmin,y = tmax, color = name)) + 
  geom_point() + 
  geom_smooth(se = FALSE) ## smooth curve occurs
```

What about the `aes` placement
```{r}
weather_df %>%
  ggplot(aes(x = tmin,y = tmax)) + 
  geom_point(aes(color = name)) + geom_smooth()
```

let's facet some things

```{r}
weather_df %>%
  ggplot(aes(x = tmin,y = tmax, color = name)) + 
  geom_point(alpha = .2) + 
  geom_smooth(se = FALSE, size = 2) +
  facet_grid(. ~ name)
  ## facet_grid(name ~ .)
```

let's combine some elements and try a new plot

```{r}
weather_df %>%
  ggplot(aes(x = date, y = tmax,color = name)) +
  geom_point(aes(size = prcp),alpha = .5) + 
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

##  some small notes

how many gemos have to exist?

You can have whatevegr geoms you want.

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax,color = name)) +
  geom_smooth(se = FALSE)
```

you can use a neat geom!

```{r}
## show how the data distribution look like
weather_df%>%
  ggplot(aes(x = tmin,y = tmax)) +
  geom_density2d() +
  geom_point(alpha = .3)
  ##geom_bin2d()  ##square shape
  ## geom_hex()
```


## univariate plots (look ad one variable at one time)

histograms are really great

```{r}
weather_df%>%
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

can we add color 

```{r}
weather_df %>%
  ggplot(aes(x = tmin,fill = name)) +
  geom_histogram(position = "dodge")
```

better: 

```{r}
weather_df %>%
  ggplot(aes(x = tmin,fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```
```{r}
weather_df %>%
  ggplot(aes(x = tmin)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

let's try a new geometry
loss something but help to compare, use adjust to be more accurate
```{r}
weather_df %>%
  ggplot(aes(x = tmin,fill = name)) +
  geom_density(alpha = .3, adjust = .5)
```

what about box plots?

```{r}
weather_df %>%
  ggplot(aes(x = name,y = tmin)) +
  geom_boxplot()
```


trendy plots

```{r}
weather_df %>%
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin() +
  stat_summary(fun = "median") ##stat_summary() shows the mean point
```

Ridge plots (popular)
better compared to the density itself
```{r}
weather_df %>%
  ggplot(aes(x = tmin,y = name)) +
  geom_density_ridges()
```

## save and embed

let's save a scatterplot

```{r}
weather_plot = 
weather_df%>%
  ggplot(aes(x = tmin,y = tmax, color = name)) +
  geom_point(alpha = .5)

ggsave("./results/weather_plot.pdf",weather_plot, width = 8, height = 5)
```

what about embedding

```{r,fig.width=6, fig.height=3}
weather_plot
```

embed at different size.


```{r, fig.width=12, fig.height=6}
weather_plot
```

```{r}
weather_df %>% 
  ggplot (aes(x = tmin, y = tmax,color = name)) +
  geom_point(alpha = .5)
```

##labels

```{r}
weather_df %>% 
  ggplot (aes(x = tmin, y = tmax,color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "minimum daily temperature(C)",
    y = "maxmium daily temperature(C)",
    caption = "Data from......"
  )
```

##scale

start with the same plot

```{r}
weather_df %>% 
  ggplot (aes(x = tmin, y = tmax,color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "minimum daily temperature(C)",
    y = "maxmium daily temperature(C)",
    caption = "Data from......"
  ) +
  scale_x_continuous(
    breaks = c(-15,0,15),
    labels = c("-15C","0","15")
  ) +
  scale_y_continuous(
    ##trans = "log"
    position = "right"
  )

```

##color scales

```{r}
weather_df %>% 
  ggplot (aes(x = tmin, y = tmax,color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "minimum daily temperature(C)",
    y = "maxmium daily temperature(C)",
    caption = "Data from......"
  ) +
  scale_color_hue(
    name = "Location",
    h = c(100,300)) ##define your own color

```


```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax,color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "minimum daily temperature(C)",
    y = "maxmium daily temperature(C)",
    caption = "Data from......"
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE)
```
## themes

shift the legend
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax,color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "minimum daily temperature(C)",
    y = "maxmium daily temperature(C)",
    caption = "Data from......"
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme(legend.position = "bottom")
```

change the overal theme

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax,color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "minimum daily temperature(C)",
    y = "maxmium daily temperature(C)",
    caption = "Data from......"
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme_minimal() +
  theme(legend.position = "bottom")
  ##theme_bw()
  ##theme_classic()
```

